EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-01 01_intro.tex
GB wide pass on the full text

Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like dynamical mean-field theory (DMFT)~\cite{Georges1996RMP}, density-matrix embedding theory (DMET)~\cite{Knizia2012_DMET} and Gutzwiller approximation (GA)~\cite{Lanata2015_GA}, where the complex behavior of the lattice systems is reduced to an effective self-consistent quantum impurity problem. The present code has been developed with a particular focus for DMFT but its usage extends also to the aforementioned methods.
}
By accounting for local quantum fluctuations, DMFT can successfully describe the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds, and high-temperature superconductors.

>>>>

Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like dynamical mean-field theory (DMFT)~\cite{Georges1996RMP}, density-matrix embedding theory (DMET)~\cite{Knizia2012_DMET} and the Gutzwiller approximation (GA)~\cite{Lanata2015_GA}, where the complex behavior of the lattice systems is reduced to an effective self-consistent quantum impurity problem. The present code has been developed with a particular focus for DMFT but its usage extends also to the aforementioned methods.
}
By accounting for local quantum fluctuations, DMFT can successfully describe the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds, and high-temperature superconductors.

### 2025-06-01 03_edipack2.tex
AA: Version 2.0 Iterated full text but intro and conclusions. Added a lot of references. Fixed (almost) all comments. Updated figAHM.
taking a pause before attacking intro  + conclusions. 

Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>
GB wide pass on the full text

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ARCMP}:
\begin{equation}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^\dagger_{\a\uparrow}d_{\a\downarrow}d^\dagger_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^\dagger_{\a\uparrow}d^\dagger_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow},\\
\end{split}
\label{Hint}
\end{equation}
where we introduced the occupation number operator $n_{\alpha\sigma}=d^\dagger_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Werner2008PRL,Werner2009PR,Mravlje2011PRL,Medici2011PRL,Medici2011PRB,Georges2013ARCMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ARCMP}. Different choices, preserving part
of the combined symmetry group, can be made for other numbers of orbitals~\cite{Georges2013ARCMP}.

@@

\begin{lstlisting}[style=fstyle,numbers=none]
!sparse row: contains bath state values for each impurity key 
type sparse_row
   integer                               :: size          !current size
   integer                               :: bath_state_min!smallest bath state
   integer                               :: bath_state_max!largest bath state
   integer,dimension(:),allocatable      :: bath_state    !Values:rank-1 bath states
   integer,dimension(:),allocatable      :: sector_indx   !bath state sector indices
end type sparse_row
!Sparse Map: array of sparse rows 
type sparse_map
   type(sparse_row),dimension(:),pointer :: imp_state     !Key: impurity states
   integer                               :: Nimp_state    !Key upper limit 
   logical                               :: status=.false.!allocation status
end type sparse_map
\end{lstlisting}

@@

For the {\bf replica} topology, the coupling between the impurity 
and each bath element is independent on spin, orbital, and internal 
bath structure: $V^p_{\a\b\s\s'} = V^p$.
In the {\bf general} case, the coupling includes explicit 
diagonal
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$. 
%
The number of independent bath parameters to optimize is either 
$N_\mathrm{params} = (N_\mathrm{sym}+1)N_\mathrm{bath}$ for {\tt bath\_type}={\bf replica} or 
$N_\mathrm{params} = (N_\mathrm{sym}+N_\a N_\s)N_\mathrm{bath}$ for {\tt bath\_type}={\bf general}.

@@

The critical functionality here is the 
determination of the total bath array dimension, implemented in 
{\tt ED\_BATH\_DIM}. The function {\tt ed\_get\_bath\_dimension} 
calculates the required dimension $N_\mathrm{params}$ for user-allocated bath 
arrays, ensuring they contain exactly the necessary number of 
parameters. 
% Any following call
% to functions hosting the bath as an input verify the array's 
% dimensionality using {\tt check\_bath\_dimension} to prevent 
% allocation errors.

The following table summarizes the total number of parameters $N_\mathrm{params}$, as returned by the function {\tt ed\_get\_bath\_dimension} and required to describe the bath according to the values of {\tt ed\_mode} and {\tt bath\_type}:

\begin{center}
\begin{tabular}{ |c||c|c|c|c| } 
 \hline     
 $N_\mathrm{params}$ & {\bf normal} & {\bf hybrid} & {\bf replica} & {\bf general}\\
 \hline 
 \hline    
 {\bf normal} & 
 $2N_\a N_\mathrm{bath}$ & 
 $(N_\a+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a)N_\mathrm{bath}$\\
 \hline
   {\bf superc} & $3N_\a N_\mathrm{bath}$ & 
 $(N_\a+2)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a)N_\mathrm{bath}$\\
 \hline
   {\bf nonsu2} & $3N_\a N_\mathrm{bath}$ & 
 $(2N_\a+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a)N_\mathrm{bath}$\\
 \hline
\end{tabular}
\end{center}

>>>>

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ARCMP}:
\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^\dagger_{\a\uparrow}d_{\a\downarrow}d^\dagger_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^\dagger_{\a\uparrow}d^\dagger_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow},\\
\end{split}
\end{equation}
where we introduced the occupation number operator $n_{\alpha\sigma}=d^\dagger_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Werner2008PRL,Werner2009PR,Mravlje2011PRL,Medici2011PRL,Medici2011PRB,Georges2013ARCMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ARCMP}. Different choices, preserving part
of the combined symmetry group, can be made for other numbers of orbitals~\cite{Georges2013ARCMP}.

@@

\begin{lstlisting}[style=fstyle,numbers=none]
!Sparse row: contains bath state values for each impurity key 
type sparse_row
   integer                               :: size          !current size
   integer                               :: bath_state_min!smallest bath state
   integer                               :: bath_state_max!largest bath state
   integer,dimension(:),allocatable      :: bath_state    !Values:rank-1 bath states
   integer,dimension(:),allocatable      :: sector_indx   !bath state sector indices
end type sparse_row
!Sparse map: array of sparse rows 
type sparse_map
   type(sparse_row),dimension(:),pointer :: imp_state     !key: impurity states
   integer                               :: Nimp_state    !key upper limit 
   logical                               :: status=.false.!allocation status
end type sparse_map
\end{lstlisting}

@@

For the {\bf replica} topology, the coupling between the impurity 
and each bath element is independent on spin, orbital, and internal 
bath structure: $V^p_{\a\b\s\s'} = V^p$.
In the {\bf general} case, the coupling includes explicit 
diagonal
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$. 
%
The number of independent bath parameters to optimize is either 
$N_\mathrm{param} = (N_\mathrm{sym}+1)N_\mathrm{bath}$ for {\tt bath\_type}={\bf replica} or 
$N_\mathrm{param} = (N_\mathrm{sym}+N_\a N_\s)N_\mathrm{bath}$ for {\tt bath\_type}={\bf general}.

@@

The critical functionality here is the 
determination of the total bath array dimension, implemented in 
{\tt ED\_BATH\_DIM}. The function {\tt ed\_get\_bath\_dimension} 
calculates the required dimension $N_\mathrm{param}$ for user-allocated bath 
arrays, ensuring they contain exactly the necessary number of 
parameters. 
% Any following call
% to functions hosting the bath as an input verify the array's 
% dimensionality using {\tt check\_bath\_dimension} to prevent 
% allocation errors.

The following table summarizes the total number of parameters $N_\mathrm{param}$, as returned by the function {\tt ed\_get\_bath\_dimension} and required to describe the bath according to the values of {\tt ed\_mode} and {\tt bath\_type}:

\begin{center}
\begin{tabular}{ |c||c|c|c|c| } 
 \hline     
 $N_\mathrm{param}$ & {\bf normal} & {\bf hybrid} & {\bf replica} & {\bf general}\\
 \hline 
 \hline    
 {\bf normal} & 
 $2N_\a N_\mathrm{bath}$ & 
 $(N_\a+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a N_\sigma)N_\mathrm{bath}$\\
 \hline
   {\bf superc} & $3N_\a N_\mathrm{bath}$ & 
 $(N_\a+2)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a N_\sigma)N_\mathrm{bath}$\\
 \hline
   {\bf nonsu2} & $3N_\a N_\mathrm{bath}$ & 
 $(2N_\a+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+1)N_\mathrm{bath}$ & 
 $(N_\mathrm{sym}+N_\a N_\sigma)N_\mathrm{bath}$\\
 \hline
\end{tabular}
\end{center}
